### 2015-03-06 paper/poly.tex
start writing down the polynomial system

\documentclass[nocopyrightspace,preprint]{sigplanconf-pldi15}

\usepackage[utf8]{inputenc} %for utf8 input
\usepackage{amssymb} %for shift symbol
\usepackage{amsmath}
\usepackage{listings} %for code
\usepackage{mathpartir} %for typing rules
\usepackage{microtype} %better micro typing
\usepackage{stmaryrd} %for llbracket
\usepackage{mathabx} % for boxes
\usepackage{graphicx} %to include png images
\usepackage{xcolor} %for colors
\usepackage{url}
\usepackage{enumitem}
\usepackage{array} %for stupid tables

%----------------------------------------------------

\usepackage{prettyref}
\newcommand{\pref}[1]{\prettyref{#1}}
\newcommand{\Pref}[1]{\prettyref{#1} \vpageref[]{#1}}
\newcommand{\ppref}[1]{\vpageref[]{#1}}
\newrefformat{fig}{Figure~\ref{#1}}
\newrefformat{app}{Appendix~\ref{#1}}
\newrefformat{tab}{Table~\ref{#1}}
\newrefformat{cha}{Challenge~\ref{#1}}
\newrefformat{compiler}{Point~\ref{#1} of \pref{thm:compiler}}

%----------------------------------------------------

\usepackage{amsthm}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}

%----------------------------------------------------

\input{definitions}
\begin{document}

\section{Canonical Polynomial Potential}

The following presentation of amortized analysis for resource
inference on integer programs permits to obtain precise polynomial
bounds by naturally extending the previous linear work. Innovations
of this presentation include
\begin{enumerate}
\item the use of canonical polynoms $X^i$ at no cost on precision,
 \\ (I wish)
\item the use of negative coefficients in potential annotations,
\item the support for operations with mixed effects on multivariate potential.
\end{enumerate}

\paragraph{Index Sets}

Let $V$ be a set of variables.  An \emph{index} $I \in \ind(V)$ is a
family that maps pairs of variables to non-zero natural numbers,
$$
I = (i_{(x,y)})_{(x,y) \in V^2} .
$$
%
We identify a family $I$ with the set of pairs $\{ ((x,y),i_{(x,y)})
\mid (x,y) \in V^2\}$.
Let $\ind(V)$ denote the set of all such indices.  We write $\ind$
instead of $\ind(V)$ if the set of variables $V$ is fixed or obvious
from the context.
%

The \emph{degree} of an index $I = (i_{(x,y)})$
is defined by
$
\deg(I) = \sum_{(x,y)} i_{(x,y)}.
$
We define $\ind_k(V) = \{ I \in \ind(V) \mid \deg(I) \leq k
\}$ to be the set of indices of degree at most $k$.  Remark
that with these notations, $\ind_0(V)$ for any $V$ is a
singleton---it contains only the empty index.

\paragraph{Resource Polynomials}

Let $V$ be a set of variabls.  An index $I \in \ind(V)$ denotes a
\emph{base polynomial} $P_I : \State \to \Q$ that maps a program
state $\state$ to a rational number.  The definition of $P_I$ is
$
P_I(\state) = \prod_{(x,y)} |[\state(x),\state(y)]|^{i_{(x,y)}}.
$
Remark that following the standard convention for $\prod$, the
empty index represents the constant base polynomial 1.

A \emph{resource polynomial} is a linear combination of
base polynomials with rational coefficients.  Resource polynomials
form a vector space over $\Q$.

\paragraph{Potential Annotations}

We call \emph{potential annotations}
the coordinates of a resource polynomial in the basis of
all base polynomials.  An annotation $Q$ is readily represented
by a collection $(q_I)_I$ binding rational numbers to indices.
For example, the potential annotation $\{\{((x,y),2)\} \mapsto 0.5,
\emptyset \mapsto 3\}$ specifies the resource polynomial
$
  P_Q(\state) = 0.5\cdot|[\state(x),\state(y)]|^2 + 3.
$

The notion of degree can be generalized to potential annotations:
the annotation $Q = (q_I)_I$ has degree $\max_{q_I} \deg(I)$.
Remark that the degree of an annotation is exactly the degree of
the polynomial it represents.

\paragraph{Potential Method for Resource Analysis}

The previous potential annotations will be used to decorate
programs and bound their resource consumption.  A program
$P$ consumes $n$ resource units turning a state $\state$ into $\state'$,
we note this $\state \rightarrow^n_P \state'$.  The \emph{potential
method} make us look for two annotations $Q$ and $Q'$ such
that
$$
  \forall n\, \state\, \state'.\, \state \rightarrow^n_P \state'
    \implies P_Q(\state) \ge n + P_{Q'}(\state').
$$
The composition of these annotations through the whole program
using a quantitative logic permits to derive a resource bound.
In the sequel we explicit a relation that is sufficient to
satisfy the previous condition when the program $P$ is an
increment operation $x \gets x+y$.

\paragraph{Shifts of Resource Polynomials}

By symmetry we can consider the case where $P=x \gets x+y$ with
$y>0$, the case where $y \le 0$ is only a few signs away.  Let
$Q$ be a potential annotation, we will describe a $Q'$ and a
$P^y_.$ map such that if $\state \rightarrow_P \state'$ (with no resource cost),
$P_Q(\state) = P^y_{Q'}(\state')$.  We call $Q'$ a shift of $Q$ and
write $Q' = \qsh_x^+ Q$.  (When $y\le0$ we get $\qsh_x^-Q$.)

The shift of $Q$ is defined on an extended set of indices, namely
$\ind^\square = \ind(V \cup \{\square\})$.  This square is not a real
variable, but a helper with special semantics.
We can now pose $\qsh_x^+ Q = (q'_I)_{I\in\ind^\square}$ where,

\begin{align*}
q'_I =
q_{I[\square \gets x]}
\prod_z
   \binom{i_{(\square,z)} {+} i_{(x,z)}}{i_{(\square,z)}}
   \binom{i_{(z,\square)} {+} i_{(z,x)}}{i_{(z,\square)}}
   (-1)^{i_{(z,\square)}}
   .
\end{align*}

The substitution operation $I[\square \gets x]$ defines an index $I'$
in $\ind(V)$ with the same degree as $I$.  This index satisfies
$$
i'_{(x,z)} {=} i_{(x,z)} {+} i_{(\square,z)}, \;
i'_{(z,x)} {=} i_{(z,x)} {+} i_{(z,\square)}, \;
i'_{(u,v)} {=} i_{(u,v)} \mbox{ o.w.}
$$
Notations may seem heavy, but what they really express is that
replacing $\square$ by $x$ in $x^2\square^3z^2$ yields $x^5z^2$.

\clearpage
\section{Binomial Polynomial Potential}

This section presents the binomial version of the previous system.
In practice, negative coefficients could be enough to avoid having to use
this basis.  I also suspect that there is no significant simplifications
in the presentation using the binomial basis.

\paragraph{Index Sets}

Let $V$ be a set of variables.  An \emph{index} $I \in \ind(V)$ is a
family that maps two-element sets of variables to natural numbers,
that is,
$$
I = (i_{\{x,y\}})_{\{x,y\} \subseteq V} \; .
$$
%
We identify a family $I$ with the set $\{ (\{x,y\},i_{\{x,y\}})
\mid \{x,y\} \subseteq V\}$.

Let $\ind(V)$ denote the set of all such indices.  We write $\ind$
instead of $\ind(V)$ if the set of variables $V$ is fixed or obvious
from the context.
%
We assume that allways $0 \in V$ and sometimes write $i_x$ instead of $i_{\{x,0\}}$.

The \emph{degree} $\deg(I)$ of an index $I = (i_{\{x,y\}})_{\{x,y\}
  \subseteq V}$ is defined as
$$
\deg(I) = \sum_{\{x,y\} \in V} i_{\{x,y\}} \;.
$$
We define $\ind_k(V) = \{ I \mid I \in \ind(V) \text{ and } \deg(i) \leq k
\}$ to be the set of indices of degree at most $k$.

\paragraph{Resource Polynomials}

Let $V$ be a set of variables.  An index $I \in \ind(V)$ denotes a
\emph{base polynomial} $P_I : \states \to \N$ for $V$ that maps a
program state $H$ to product of binomial coefficients (a natural
number).  We define
$$
P_I(H) = \prod_{{\{x,y\}} \subseteq V} \binom {|H(x){-}H(y)|} {i_{\{x,y\}}} \; .
$$
%
A \emph{resource polynomial} $R$ for the variable set $V$ is a
non-negative linear combination of the base polynomials for $V$.

\paragraph{Potential Annotations}

A \emph{potential annotation} for the variable set $V$ is a family
$$Q = (q_I)_{I \in \ind(V)}$$
of non-negative rational numbers.  Such an annotation denotes the
resource polynomial $R_Q$ that is defined by
$$
R_Q(H) = \sum_{I \in \ind(V)} q_I \cdot P_I(H) \; .
$$
%
We say that the potential annotation $Q$ is of degree $k$ if $q_I = 0$
for $I \in \ind(V)$ with $\deg(I) > k$.

\paragraph{Additive Shifts}

Let $Q$ be a potential annotation for a variable set $V$ and let
$\{x,y\} \subseteq V$ be a two-element variable set.  The
\emph{additive shift} with respect to $\{x,y\}$ is a potential
annotation $\shift_{\{x,y\}}(Q) = (q'_I)_{I \in \ind(V)} $ for $V$
that is defined through
$$
q'_I = q_I + q_{I^{\{x,y\}{+}1}} \; .
$$
For an index $I = (i_{\{x,y\}})_{\{x,y\} \subseteq V}$ we use the
notation $I^{\{x,y\}{+}k}$ to denote the index
$(i'_{\{x,y\}})_{\{x,y\} \subseteq V}$ such that
$$
i'_{\{t,u\}} = \left\{
  \begin{array}{ll}
    i_{\{t,u\}} + k  & \text{if } \{t,u\} = \{x,y\} \\
    i_{\{t,u\}} & \text{otherwise}
  \end{array}
\right.
\;.
$$
%
The additive shift for natural numbers reflects the identity
\begin{equation}
\label{eq:shift}
\sum_{0 {\leq} i \leq {k}} q_i \binom{n+1}{i} = \sum_{0 {\leq} i \leq {k}} (q_i{+}q_{i+1}) \binom{n}{i}
\end{equation}
where $q_{k+1} = 0$.  It is used in the effect system if the
difference $n+1$ between two variables $x,y$ decreases by one.

\begin{lemma} Let $V$ be a set of variables with $x,y \in V$ and let
  $H$ be a program state. Let $|H'(t) {-} H'(u)| = |H(t) {-} H(u)|$
  for $\{t,u\} \neq \{x,y\}$ and let $|H'(x) {-} H'(y)| = |H(x) {-}
  H(y)| - 1$.
  %
  If $Q' = \shift_{\{x,y\}}(Q)$ then $R_Q(H) = R_{Q'}(H')$.
\end{lemma}

We now study the effect of multiple simultaneous shifts.  Let $Q$ be a
resource annotation for a variable set $V$ and let $U_1,\ldots,U_n
\subseteq V$ with $|U_i| = 2$ for all $i$ and $U_i \neq U_j$ for $i
\neq j$ be pairwise distinct two-element variable sets.  The
simulations additive shift $\shift_{U_1,\ldots,U_n}(Q)$ of $Q$ with
respect to $U_1,\ldots,U_n$ is defined by
$$
\shift_{U_1,\ldots,U_n}(Q) = \shift_{U_1}( \cdots \shift_{U_n}(Q) \cdots ) \; .
$$
%
\begin{proposition}
  Let $V$ be a set of variables and let $U_1,\ldots,U_n$ be pairwise
  distinct two-element variable sets.  Let $|H'(x) {-} H'(y)| = |H(x)
  {-} H(y)|$ for $\{x,y\} \not\in \{U_1,\ldots,U_n\}$ and let $|H'(x)
  {-} H'(y)| = |H(x) {-} H(y)| - 1$ for $\{x,y\} \in
  \{U_1,\ldots,U_n\}$.
  %
  If $Q' = \shift_{\{U_1,\ldots,U_n\}}(Q)$ then $R_Q(H) = R_{Q'}(H')$.
\end{proposition}
%
As shown by the following lemma, the order in which the shifts for the
individual $U_i$ are applied is insignificant.
%
\begin{lemma}
  Let $\sigma : \{1,\ldots,n\} \to \{1,\ldots,n\}$ be a
  permutation. Then $\shift_{U_1,\ldots,U_n}(Q) =
  \shift_{U_{\sigma(1)},\ldots,U_{\sigma(n)}}(Q)$.
\end{lemma}
%
For reasons of efficiency in the constraint generation, we give a more
direct formula for the simultaneous shift.  Let $I \in \ind(V)$ and
let $U_1,\ldots,U_n$ be pairwise distinct two-element variable sets.
We define the index $I^{U_1,\ldots,U_n + k}$ as the family $(i'_{\{x,y\}})_{\{x,y\} \subseteq V}$ such that
$$
i'_{\{t,u\}} = \left\{
  \begin{array}{ll}
    i_{\{t,u\}} + k  & \text{if } \{t,u\} \in \{U_1,\ldots,U_n\} \\
    i_{\{t,u\}} & \text{otherwise}
  \end{array}
\right.  \;.
$$

%
\begin{lemma}
  Let $V$ be a variable set and let $U_1,\ldots,U_n$ be pairwise
  distinct two-element variable sets.
  %
  Let $Q = (q_I)_{I \in \ind(V)}$ be a resource annotation for
  $V$ and let $ Q' = (q'_I)_{I \in \ind(V)}$ where
  $$
  q'_I = \sum_{\{j_1,\ldots,j_m\} \subseteq \{1,\ldots,n\} } q_{I^{U_{j_1},\ldots,U_{j_m}+1}} \; .
  $$
  Then $Q' = \shift_{U_1,\ldots,U_n}(Q)$.
\end{lemma}

\subsection{Binomial basis}

\begin{lemma}
  There exists a family $(c^{ij}_k)_k$, such that
  $$
  \sum_k c^{ij}_k \binom n k = \binom n i \binom n j,
  $$
  and $c^{ij}_k = \binom k j \binom j {i+j - k}$ satisfies
  the previous equation.
\end{lemma}
\begin{proof}
  \begin{align*}
    \binom n i \binom n j
      &= \sum_k \binom {n-j} k \binom j {i-k} \binom n j \\
      &= \sum_k \binom {n-j} {k-j} \binom n j \binom j {i+j-k} \\
      &= \sum_k \binom n k \binom k j \binom j {i+j-k}.
  \end{align*}
  In the first equality, we use Vandermonde's convolution
  formula on $\binom n i$.  In the second equality we change
  the summation variable to $k-j$.  And in the third equation
  we use the identity proved below.
  \begin{align*}
    \binom {n-j} {k-j} \binom n j
      &= \frac {(n-j)!} {(n-k)! (k-j)!} \; \frac {n!} {(n-j)! j!} \\
      &= \frac {n!} {(n-k)! (k-j)! j!} \\
      &= \frac {n!} {(n-k)! k!} \; \frac {k!} {(k-j)! j!} \\
      &= \binom n k \binom k j.
  \end{align*}
\end{proof}

%%
% Note: Later we need to shift in many directions at once like
%   Q' = shift_{x,y} (shift_{x,u} (Q))
% To do: Give a combined formula for that (concise constraint system).
%%

\end{document}